How Optimal is Word Recognition Under Multimodal Uncertainty?

### 2019-11-14 Writeup/ms_r2.tex
final

\linenumbers
\usepackage[sortcites=false,sorting=none]{biblatex}

@@

Language uses symbols expressed in one modality --- the auditory
modality, in the case of speech --- to communicate about the world,
which we perceive through many different sensory modalities. Consider
hearing someone yell \enquote{bee!} at a picnic, as a honey bee buzzes
around the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence.

However, word identification takes place in a noisy world, and the cues
received through each modality may not provide a definitive answer. On
the auditory side, individual acoustic word tokens are almost always
ambiguous with respect to the particular sequence of phonemes they
represent, which is due to the inherent variability of how a phonetic
category is realized acoustically (Hillenbrand, Getty, Clark, \&
Wheeler, 1995). And some tokens may be distorted additionally by
mispronunciation or ambient noise. Perhaps the speaker was yelling
\enquote{pea} and not \enquote{bee.} Similarly, a sensory impression may
not be enough to make a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

As a simplified case study of early word learning, the task of matching
sounds to corresponding visual objects has been studied extensively in
the developmental literature. For example, many studies focus on how
children might succeed in this type of task despite referential
ambiguity (Medina, Snedeker, Trueswell, \& Gleitman, 2011; Pinker, 1989;
Smith \& Yu, 2008; Suanda, Mugwanya, \& Namy, 2014; Vlach \& Johnson,
2013; Vouloumanos, 2008; Yurovsky \& Frank, 2015). However, even when
they have learned the exact meaning of a word, observers (both children
and adults) often still find it challenging to recognize which word the
speaker has uttered, especially under noise (Mattys, Davis, Bradlow, \&
Scott, 2012; Peelle, 2018). The purpose of the current study is thus to
explore word recognition by adults under multimodal uncertainty,
focusing on the special case where people have access to multimodal cues
from the auditory speech and the visual referent. In the General
Discussion, we return to the question of how these findings relate to
questions about word learning.

One rigorous way to approach this question is through conducting an
\emph{ideal observer} analysis. This research strategy provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization.\footnote{It is, thus, a general
  instance of the rational approach to cognition (Anderson, 1990),
  instantiating Marr's computational level of analysis (Marr, 1982).}
When there is uncertainty in the input, the ideal observer performs an
optimal probabilistic inference. For example, in order to recognize an
ambiguous linguistic input, the model uses all available probabilistic
knowledge in order to maximize the accuracy of this recognition. The
ideal observer model can be seen as a theoretical upper limit on
performance. It is not so much a realistic model of human performance,
as much as a baseline against which human performance can be compared
(Geisler, 2003; Rahnev \& Denison, 2018). When there is a deviation from
the ideal, it can reveal extra constraints on human cognition, such as
limitations on the working memory or attentional resources. This
approach has had a tremendous impact not only on speech-related research
(Clayards, Tanenhaus, Aslin, \& Jacobs, 2008; Feldman, Griffiths, \&
Morgan, 2009; Kleinschmidt \& Jaeger, 2015; Norris \& McQueen, 2008),
but also on many other disciplines in the cognitive sciences (for
reviews, see Chater \& Manning, 2006; Knill \& Pouget, 2004; Tenenbaum,
Kemp, Griffiths, \& Goodman, 2011)

Some prior ideal observer studies are closely related to the question we
are addressing in the current work. For instance, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (Voice Onset Time) that differentiated similar
words (e.g., \enquote{beach} and \enquote{peach}). They found that
humans were sensitive to these probabilistic cues and their judgments
closely reflected the optimal predictions. And Feldman et al. (2009)
studied the perceptual magnet effect, a phenomenon that involves reduced
discriminability near prototypical sounds in the native language (Kuhl,
1991), showing that this effect can be explained as the consequence of
optimally solving the problem of perception under uncertainty (see also
Kronrod, Coppess, \& Feldman, 2016).

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the predictions of an ideal observer. This
previous research did not, however, study speech understanding when
visual information was obtained through the referential context rather
than through observation of speaker's face. Although some experimental
findings show that information about the identity of a referent can be
integrated with linguistic information to resolve lexical and syntactic
ambiguities in speech (e.g., Eberhard, Spivey-Knowlton, Sedivy, \&
Tanenhaus, 1995; Spivey, Tanenhaus, Eberhard, \& Sedivy, 2002;
Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995), to our knowledge
no study has offered an ideal observer analysis of this task.

@@

First, in the case of audio-visual speech, both modalities offer
information about the same underlying speech category. They differ only
in terms of their informational reliability. In a referential context,
however, the auditory and visual modalities play different roles in the
referential process -- the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning} (and these
differences are in addition to possible differences in informational
reliability). Further, speech is claimed to have a privileged status
compared to other sensory stimuli (Edmiston \& Lupyan, 2015; Lupyan \&
Thompson-Schill, 2012; Vouloumanos \& Waxman, 2014; Waxman \& Gelman,
2009; Waxman \& Markow, 1995), and this privilege is suggested to be
specifically related to the ability to refer (Waxman \& Gelman,
2009).\footnote{There is, however, a debate as to whether speech is
  privileged for children and adults for similar reasons. Whereas some
  researchers suggest that speech is privileged for both children and
  adults because of its ability to refer (e.g., Waxman \& Gelman, 2009),
  others suggest that speech might \emph{not} have a referential status
  from the start. Rather, speech might be preferred by children only
  because of a low level auditory ``overshadowing'' (e.g., Sloutsky \&
  Napolitano, 2003).} Thus, in a referential context, it is possible
that listeners do not treat the auditory and visual modalities as
equivalent sources of information. Instead, there could be a sub-optimal
bias for the auditory modality beyond what is expected from
informational reliability alone.

@@

In the testing phase of the original task (Sloutsky \& Napolitano,
2003), participants were asked whether or not the two audio-visual
presentations are \emph{identical}. In the current study, we are
interested, rather, in the categorization, i.e., determining whether or
not two similar tokens are members of the same phonological/semantic
category. Therefore, testing in our task is category-based: Participants
are asked to press \enquote{same} if they think the second item (the
test) belongs to the same category as the first (target) (e.g.,
dog-/aba/), even if there is a slight difference in the sound, in the
referent, or in both. They are instructed to press \enquote{different}
only if they think that the second stimulus was an instance of the other
category (cat-/ada/). The task also includes trials where pictures are
hidden (audio-only) or where sounds are muted (visual-only). These
unimodal trials provide us with the participants' evaluation of the
probabilistic information present in the auditory and visual categories.
As we shall see, these unimodal distributions are used as inputs to the
optimal cue combination model.

@@

We assume that both the auditory categories (i.e., /aba/ and /ada/) and
the visual categories (cat and dog) are distributed along a single
acoustic and semantic dimension, respectively (Figure~\ref{fig:model}).
Moreover, we assume that all categories are normally distributed.
Formally speaking, if \(A\) denotes an auditory category (/ada/ or
/aba/), then the probability that a point \(a\) along the acoustic
dimension belongs to the category \(A\) is
\[ p(a | A) \sim  N(\mu_A, \sigma^2_A) \] where \(\mu_A\) and
\(\sigma^2_A\) are respectively the mean and the variance of the
auditory category. Similarly, the probability that a point \(v\) along
the visual dimension belongs to the category \(V\) is
\[ p(v | V) \sim  N(\mu_V, \sigma^2_V) \] where \(\mu_V\) and
\(\sigma^2_V\) are the mean and the variance of the visual category. An
audio-visual signal \(w=(a,v)\) can be represented as a point in the
audio-visual space. These audio-visual tokens define bivariate
distributions in the bi-dimentional space. We call these bivariate
distributions \emph{Word categories}, noted \(W\), and are distributed
as follows: \[ p(w | W) \sim  N(M_W, \Sigma_W) \] where
\(M_W=(\mu_A, \mu_V)\) and \(\Sigma_W\) are the mean and the covariance
matrix of the word category. The main assumption of the model is that
the auditory and visual variables are independent (i.e., uncorrelated),
so the covariance matrix is simply: \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

@@

\noindent This assumption says that, given a word-object mapping, e.g.,
\(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid assumption in the context of our
task.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

@@

The parameter \(b\) represents the differential between the categories'
prior probabilities. However, since the identity of word categories is
randomized across participants, \(b\) measures, rather, a response bias
to \enquote{same} if \(b > 0\), and a response bias to
\enquote{different} if \(b < 0\). We expect a general bias towards
answering \enquote{different} because of the categorical nature of our
same-different task: When two items are ambiguous but perceptually
different, participants might have a slight preference for
\enquote{different} over \enquote{same}. As for the means, their values
are fixed, and they correspond to the most typical tokens in our
stimuli. Finally, observations from each modality (\(a\) and \(v\)) are
weighted in Equation 1 according to their reliability (that is,
according to the \emph{inverse} of their variance):
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}.\]

@@

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their parameters (i.e., variances associated
with the visual and auditory modalities) are derived from data in the
unimodal conditions. In addition to these normative models, we consider
a \emph{descriptive} model. It is formally identical to the normative
optimal model (Equation 1), except that the parameters are fit to actual
responses in the bimodal condition. If the referential task induces
sub-optimality (due, for instance, to the arbitrary nature of the
sound-object association), then the descriptive model should explain
more variance than the optimal model does.

Comparison of the optimal and the descriptive models allows us, not only
to quantify how much people deviate from optimality, but also to
understand precisely the nature of this deviation. Let \(\sigma^2_{A}\)
and \(\sigma^2_{V}\) be the values of the variances used in the optimal
model (derived from the unimodal conditions), and \(\sigma^2_{Ab}\) and
\(\sigma^2_{Vb}\) be the values observed through the descriptive model
in the bimodal condition. Deviation from optimality is measured in two
ways. First, we measure the change in the values of the variance
specific to each modality, that is, how \(\sigma^2_{A}\) compares to
\(\sigma^2_{Ab}\), and how \(\sigma^2_{V}\) compares to
\(\sigma^2_{Vb}\). Second, we measure changes in the proportion of the
visual and auditory variances, i.e., we examine how
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\) compares to
\(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). The first measure allows us to
test if response precision changes for each modality when we move from
the unimodal to the bimodal conditions. The second allows us to test the
extent to which the weighting scheme follows the prediction of the
optimal model. The reason we used the proportion of the variances as a
measure of cross-modal weighting is because this proportion corresponds
to the
slope\footnote{Or more precisely the absolute value of the slope.} of
the decision threshold in the audio-visual space
(Figure~\ref{fig:model}). The decision threshold is defined as the set
of values in this audio-visual space along which the posterior is equal
to 0.5. Formally speaking, the decision threshold has the following
form:

@@

In this Experiment, we test the predictions of the model in the case
where uncertainty is due to categorical variability (i.e., ambiguity in
terms of category membership) and inherent sensory noise. We do not add
any external noise to the background. Thus, we test the following
(normative) cue weighting scheme:

@@

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. Participants were excluded if
they reported having experienced a technical problem of any sort during
the online experiment (N=14), or if they had less than 50\% accurate
responses on the unambiguous training trials (N=6). The final sample
consisted of N = 80 participants. All participants provided informed
consent before taking the experiment.
\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

@@

Remember that data in these conditions allows us to derive the variances
of both the auditory and the visual categories, and that these variances
are used to make predictions about bimodal data (in the visual and
auditory baselines as well as in the optimal model). These individual
variances were derived as follows (we explain the derivation for the
auditory-only case, but the same applies for the visual-only case). We
use the same Bayesian reasoning as we did in the derivation of the
bimodal model: When presented with an audio instance \(a\), the
probability of choosing the sound category 2 (that is, to answer
\enquote{different}) is the posterior probability of this category
\(p(A_2|a)\). If we assume that both sound categories have equal
variances, the posterior probability reduces to:

@@

In this condition, only the descriptive model is fit to the data, using
the expression of the posterior (Equation 1). Since the values of the
means are fixed, we have 3 free parameters: the variances for the visual
and the auditory modalities, respectively, and \(b\), the response bias.
The visual and auditory baselines as well as the optimal model are not
fit to the bimodal data, but their predictions are tested against these
bimodal data. All these normative models use the variances derived from
the unimodal data and the bias term derived from the fit to bimodal
data.

@@

\caption{\label{tab:param}Statistics for the dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Auditory} & \multicolumn{2}{c}{Visual} & \multicolumn{3}{c}{Bimodal} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5} \cmidrule(l{2pt}r{2pt}){6-8}
Experiment & b\textsubscript{A} & Var\textsubscript{A} & b\textsubscript{V} & Var\textsubscript{V} & b\textsubscript{b} & Var\textsubscript{Ab} & Var\textsubscript{Vb}\\
\midrule
Experiment1 & -0.20 & 2.04 & -0.12 & 3.33 & -0.34 & 4.96 & 7.06\\
Experiment2 & -0.18 & 4.70 & -0.24 & 3.93 & -0.38 & 9.84 & 5.21\\
Experiment3 & -0.24 & 1.94 & 0.11 & 13.00 & -0.35 & 3.00 & 39.42\\
Experiment4 & -0.40 & 1.92 & -0.22 & 3.24 & -0.42 & 4.17 & 7.28\\
\bottomrule
\end{tabular}
\end{table}

@@

Average categorization judgments and best fits are shown in Figure
\ref{fig:unimodal}. The categorization function of the auditory
condition was slightly steeper than that of the visual condition,
meaning that participants perceived the sound tokens slightly more
categorically and with higher certainty than they did with the visual
tokens. The unimodal models' estimates are shown in Table
\ref{tab:param}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/unimodal-1} \caption{Human responses in the unimodal conditions across the three experiments. Points represent the proportion of `different' to `same' responses in the auditory-only condition (left), and visual-only condition (right). Error bars are 95\% confidence intervals. Solid lines represent best unimodal posterior fits.}\label{fig:unimodal}
\end{figure}

@@

Figure \ref{fig:bimodal} compares the predictions of the normative and
descriptive models against human responses. Remember that the normative
models use the parameters estimated from the unimodal conditions (where
people see input from only one modality) to predict behavior in the
bimodal condition (where people see input from both modalities). The
descriptive model has a similar structure than the optimal model, but is
directly fit to human responses in the bimodal condition in order to
allow us to assess deviation from optimality.

We found, through comparing the correlation values, that the optimal
model explained more variance than the visual and auditory models did.
However, the optimal model was not perfect: It explained less variance
than the descriptive model did, which indicates a deviation from
optimality. To investigate this deviation, we compare the parameter
values of the optimal model to the values obtained in the descriptive
model (Table
\ref{tab:param}).\footnote{Note that the descriptive model explained almost all the variance in mean responses, which makes it a reasonable proxy for human real performance in the bimodal condition.}.
We note an increase in \emph{both} the auditory and visual variances.
This increase in noise is compatible with the fact that human responses
appear to be pulled towards chance (i.e., the value 0.5) when compared
to the optimal model (see \ref{fig:bimodal}). Below we investigate if
this deviation from optimality can be related to the cue combination
strategy.

@@

A sample of 100 participants was recruited online through Amazon
Mechanical Turk. We used the same exclusion criteria as in Experiment 1.
7 participants were excluded because they had less than 50\% accurate
responses on the unambiguous training trials. The final sample consisted
of N = 93 participants.

@@

>>>>

\linenumbers
\usepackage{xcolor}
\usepackage[sortcites=false,sorting=none]{biblatex}

@@

Language uses symbols expressed in one modality --- the auditory
modality, in the case of speech --- to communicate about the world,
which we perceive through many different sensory modalities. Consider
hearing someone yell \enquote{bee!} at a picnic, as a honey bee buzzes
around the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provides convergent evidence.

However, word identification takes place in a noisy world, and the cues
received through each modality may not provide a definitive answer. On
the auditory side, individual acoustic word tokens are almost always
ambiguous with respect to the particular sequence of phonemes they
represent, which is due to the inherent variability of how a phonetic
category is realized acoustically (Hillenbrand, Getty, Clark, \&
Wheeler, 1995). Moreover, some tokens may be distorted additionally by
mispronunciation or ambient noise. Perhaps the speaker was yelling
\enquote{pea} and not \enquote{bee.} Similarly, a sensory impression may
not be enough to make a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be accomplished through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

As a simplified case study of early word learning, the task of matching
sounds to corresponding visual objects has been studied extensively in
the developmental literature. For example, many studies focus on how
children might succeed in this type of task despite referential
ambiguity (Medina, Snedeker, Trueswell, \& Gleitman, 2011; Pinker, 1989;
L. B. Smith \& Yu, 2008; Suanda, Mugwanya, \& Namy, 2014; Vlach \&
Johnson, 2013; Vouloumanos, 2008; Yurovsky \& Frank, 2015). However,
even when they have learned the exact meaning of a word, observers (both
children and adults) often still find it challenging to recognize which
word the speaker has uttered, especially under noise (Mattys, Davis,
Bradlow, \& Scott, 2012; Peelle, 2018). The purpose of the current study
is thus to explore word recognition by adults under multimodal
uncertainty, focusing on the special case where people have access to
multimodal cues from the auditory speech and the visual referent. In the
General Discussion, we return to the question of how these findings
relate to questions about word learning.

One rigorous way to approach this question is through conducting an
\emph{ideal observer} analysis. This research strategy provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization.\footnote{It is, thus, a general
  instance of the rational approach to cognition (Anderson, 1990),
  instantiating Marr's computational level of analysis (Marr, 1982).}
When there is uncertainty in the input, the ideal observer performs an
optimal probabilistic inference. For example, in order to recognize an
ambiguous linguistic input, the model uses all available probabilistic
knowledge in order to maximize the accuracy of this recognition. The
ideal observer model can be seen as a theoretical upper limit on
performance. It is not so much a realistic model of human performance,
as much as a baseline against which human performance can be compared
(Geisler, 2003; Rahnev \& Denison, 2018). When there is a deviation from
the ideal, it can reveal extra constraints on human cognition, such as
limitations on the working memory or on attentional resources. This
approach has had a tremendous impact not only on speech-related research
(Clayards, Tanenhaus, Aslin, \& Jacobs, 2008; Feldman, Griffiths, \&
Morgan, 2009; Kleinschmidt \& Jaeger, 2015; Norris \& McQueen, 2008) but
also on many other disciplines in the cognitive sciences (for reviews,
see Chater \& Manning, 2006; Knill \& Pouget, 2004; Tenenbaum, Kemp,
Griffiths, \& Goodman, 2011).

Some prior ideal observer studies are closely related to the question we
are addressing in the current work. For instance, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (Voice Onset Time) that differentiated similar
words (e.g., \enquote{beach} and \enquote{peach}). They found that
humans were sensitive to these probabilistic cues and their judgments
closely reflected the optimal predictions. Moreover, Feldman et al.
(2009) studied the perceptual magnet effect, a phenomenon that involves
reduced discriminability near prototypical sounds in the native language
(Kuhl, 1991), showing that this effect can be explained as the
consequence of optimally solving the problem of perception under
uncertainty (see also Kronrod, Coppess, \& Feldman, 2016).

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the predictions of an ideal observer. This
previous research did not, however, study speech understanding when
visual information was obtained through the referential context rather
than through observation of the speaker's face. Although some
experimental findings show that information about the identity of a
referent can be integrated with linguistic information to resolve
lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995), to our knowledge no study has offered an ideal observer
analysis of this task (as we do here).

@@

First, in the case of audio-visual speech, both modalities offer
information about the same underlying speech category. They differ only
in terms of their informational reliability. In a referential context,
however, the auditory and visual modalities play different roles in the
referential process --- the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning} (and these
differences are in addition to possible differences in informational
reliability). Speech is claimed to have a privileged status compared to
other sensory stimuli (Edmiston \& Lupyan, 2015; Lupyan \&
Thompson-Schill, 2012; Vouloumanos \& Waxman, 2014; Waxman \& Gelman,
2009; Waxman \& Markow, 1995), and this privilege is suggested to be
specifically related to the ability to refer (Waxman \& Gelman,
2009).\footnote{There is, however, a debate as to whether speech is
  privileged for children and adults for similar reasons. Whereas some
  researchers suggest that speech is privileged for both children and
  adults because of its ability to refer (e.g., Waxman \& Gelman, 2009),
  others suggest that speech might \emph{not} have a referential status
  from the start. Rather, speech might be preferred by children only
  because of a low level auditory ``overshadowing'' (e.g., Sloutsky \&
  Napolitano, 2003). We do not take a stance on this issue here.} Thus,
in a referential context, it is possible that listeners do not treat the
auditory and visual modalities as equivalent sources of information.
Instead, there could be a (potentially sub-optimal) bias for the
auditory modality beyond what is expected from informational reliability
alone.

@@

In the testing phase of the original task (Sloutsky \& Napolitano,
2003), participants were asked whether or not the two audio-visual
presentations are \emph{identical}. In the current study, we are
interested, rather, in the categorization, i.e., determining whether or
not two similar tokens are members of the same phonological/semantic
category. Therefore, testing in our task is category-based: Participants
are asked to press \enquote{same} if they think the second item (the
test) belongs to the same category as the first (target) (e.g.,
cat-/aba/), even if there is a slight difference in the sound, in the
referent, or in both. They are instructed to press \enquote{different}
only if they think that the second stimulus was an instance of the other
category (cat-/ada/). The task also includes trials where pictures are
hidden (audio-only) or where sounds are muted (visual-only). These
unimodal trials provide us with the participants' evaluation of the
probabilistic information present in the auditory and visual categories.
As we shall see, these unimodal distributions are used as inputs to the
optimal cue combination model.

@@

We assume that both the auditory categories (i.e., /aba/ and /ada/) and
the visual categories (cat and dog) are distributed along a single
acoustic and semantic dimension, respectively (Figure~\ref{fig:model}).
Moreover, we assume that all categories are normally distributed.
Formally speaking, if \(A\) denotes an auditory category (/ada/ or
/aba/), then the probability that a point \(a\) along the acoustic
dimension belongs to the category \(A\) is
\[ p(a | A) \sim  N(\mu_A, \sigma^2_A) \] where \(\mu_A\) and
\(\sigma^2_A\) are respectively the mean and the variance of the
auditory category. Similarly, the probability that a point \(v\) along
the visual dimension belongs to the category \(V\) is
\[ p(v | V) \sim  N(\mu_V, \sigma^2_V) \] where \(\mu_V\) and
\(\sigma^2_V\) are the mean and the variance of the visual category. An
audio-visual signal \(w=(a,v)\) can be represented as a point in the
audio-visual space. These audio-visual tokens define bivariate
distributions in the bi-dimensional space. We call these bivariate
distributions \emph{Word categories}, noted \(W\), and are distributed
as follows: \[ p(w | W) \sim  N(M_W, \Sigma_W) \] where
\(M_W=(\mu_A, \mu_V)\) and \(\Sigma_W\) are the mean and the covariance
matrix of the word category. The main assumption of the model is that
the auditory and visual variables are independent (i.e., uncorrelated),
so the covariance matrix is simply: \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

@@

\noindent This assumption says that, given a word-object mapping, e.g.,
\(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid assumption in the context of our
task.\footnote{Note that this assumption is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

@@

The parameter \(b\) represents the differential between the categories'
prior probabilities. However, since the identity of word categories is
randomized across participants, \(b\) measures, rather, a response bias
to \enquote{same} if \(b > 0\), and a response bias to
\enquote{different} if \(b < 0\). We expect a general bias towards
answering \enquote{different} because of the categorical nature of our
same-different task: When two items are ambiguous but perceptually
different, participants might have a slight preference for
\enquote{different} over \enquote{same.} As for the means, their values
are fixed, and they correspond to the most typical tokens in our
stimuli. Finally, observations from each modality (\(a\) and \(v\)) are
weighted in Equation 1 according to their reliability (that is,
according to the \emph{inverse} of their variance):
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}.\]

@@

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. It is formally identical to the
normative optimal model (Equation 1), except that the parameters are fit
to actual responses in the bimodal condition. If the referential task
induces sub-optimality (due, for instance, to the arbitrary nature of
the sound-object association), then the descriptive model should explain
more variance than the optimal model does.

Comparison of the optimal and the descriptive models allows us, not only
to quantify how much people deviate from optimality but also to
understand precisely the nature of this deviation. Let \(\sigma^2_{A}\)
and \(\sigma^2_{V}\) be the values of the variances used in the optimal
model (derived from the unimodal conditions), and \(\sigma^2_{Ab}\) and
\(\sigma^2_{Vb}\) be the values observed through the descriptive model
in the bimodal condition. Deviation from optimality is measured in two
ways. First, we measure the change in the values of the variance
specific to each modality, that is, how \(\sigma^2_{A}\) compares to
\(\sigma^2_{Ab}\), and how \(\sigma^2_{V}\) compares to
\(\sigma^2_{Vb}\). Second, we measure changes in the proportion of the
visual and auditory variances, i.e., we examine how
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\) compares to
\(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). The first measure allows us to
test if response precision changes for each modality when we move from
the unimodal to the bimodal conditions. The second allows us to test the
extent to which the weighting scheme follows the prediction of the
optimal model. The reason we used the proportion of the variances as a
measure of cross-modal weighting is that this proportion corresponds to
the slope\footnote{Or more precisely the absolute value of the slope.}
of the decision threshold in the audio-visual space
(Figure~\ref{fig:model}). The decision threshold is defined as the set
of values in this audio-visual space along which the posterior is equal
to 0.5. Formally speaking, the decision threshold has the following
form:

@@

In this Experiment, we test the predictions of the model in the case
where uncertainty is due to categorical variability (i.e., ambiguity in
terms of category membership) and inherent sensory noise. We do not add
any external noise to the background. Thus, we test the following
(normative) cue weighting scheme.\footnote{In all our analyses, we study
  the combined effect of both categorical and sensory uncertainties in
  the process of cue combination. We do not examine the specific role of
  each source of uncertainty separately. For example, in the auditory
  modality, the quantity \(\sigma^2_{A_C} + \sigma^2_{A_N}\) will be
  treated as a single variable. This variable will be measured in the
  unimodal condition and tested in the bimodal condition.}

@@

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. Participants were excluded if
they reported having experienced a technical problem of any sort during
the online experiment (\(N =\) 14), or if they had less than 50\%
accurate responses on the unambiguous training trials (\(N =\) 6). The
final sample consisted of \(N =\) 80 participants. All participants
provided informed consent before taking the experiment.
\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

@@

Remember that data in these conditions allows us to derive the variances
of both the auditory and the visual categories and that these variances
are used to make predictions about bimodal data (in the visual and
auditory baselines as well as in the optimal model). These individual
variances were derived as follows (we explain the derivation for the
auditory-only case, but the same applies to the visual-only case). We
use the same Bayesian reasoning as we did in the derivation of the
bimodal model: When presented with an audio instance \(a\), the
probability of choosing the sound category 2 (that is, to answer
\enquote{different}) is the posterior probability of this category
\(p(A_2|a)\). If we assume that both sound categories have equal
variances, the posterior probability reduces to:

@@

In this condition, only the descriptive model is fit to the data, using
the expression of the posterior (Equation 1). Since the values of the
means are fixed, we have 3 free parameters: the variances for the visual
and the auditory modalities, respectively, and \(b\), the response bias.
The visual and auditory baselines, as well as the optimal model, are not
fit to the bimodal data, but their predictions are tested against these
bimodal data. All these normative models use the variances derived from
the unimodal data and the bias term derived from the fit to bimodal
data.

@@

\caption{\label{tab:param}Estimates of the models' parameters.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Auditory} & \multicolumn{2}{c}{Visual} & \multicolumn{3}{c}{Bimodal} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5} \cmidrule(l{2pt}r{2pt}){6-8}
Experiment & b\textsubscript{A} & Var\textsubscript{A} & b\textsubscript{V} & Var\textsubscript{V} & b\textsubscript{b} & Var\textsubscript{Ab} & Var\textsubscript{Vb}\\
\midrule
Experiment1 & -0.20 & 2.04 & -0.12 & 3.33 & -0.34 & 4.96 & 7.06\\
Experiment2 & -0.18 & 4.70 & -0.24 & 3.93 & -0.38 & 9.84 & 5.21\\
Experiment3 & -0.24 & 1.94 & 0.11 & 13.00 & -0.35 & 3.00 & 39.42\\
Experiment4 & -0.40 & 1.92 & -0.22 & 3.24 & -0.42 & 4.17 & 7.28\\
\bottomrule
\end{tabular}
\end{table}

@@

Average categorization judgments and best fits are shown in Figure
\ref{fig:unimodal}. The categorization function of the auditory
condition was slightly steeper than that of the visual condition,
meaning that participants perceived the sound tokens slightly more
categorically and with higher certainty than they did with the visual
tokens. The unimodal models' estimates in each modality are shown in
Table \ref{tab:param}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_r2_files/figure-latex/unimodal-1} \caption{Human responses in the unimodal conditions across the four experiments. Points represent the proportion of `different' to `same' responses in the auditory-only condition (left), and visual-only condition (right). Error bars are 95\% confidence intervals. Solid lines represent best unimodal posterior fits.}\label{fig:unimodal}
\end{figure}

@@

Figure \ref{fig:bimodal} compares the predictions of the normative and
descriptive models against human responses. Remember that the normative
models use the parameters estimated from the unimodal conditions (where
people see input from only one modality) to predict behavior in the
bimodal condition (where people see input from both modalities). The
descriptive model has a similar structure than the optimal model but is
directly fit to human responses in the bimodal condition in order to
allow us to assess deviation from optimality.

We found, by comparing the correlation values, that the optimal model
explained more variance than the visual and auditory models did.
However, the optimal model was not perfect: It explained less variance
than the descriptive model did, which indicates a deviation from
optimality. To investigate this deviation, we compare the parameter
values of the optimal model to the values obtained in the descriptive
model (Table
\ref{tab:param}).\footnote{Note that the descriptive model explained almost all the variance in mean responses, which makes it a reasonable proxy for human real performance in the bimodal condition.}
We note an increase in \emph{both} the auditory and visual variances.
This increase in noise is compatible with the fact that human responses
appear to be pulled towards chance (i.e., the value 0.5) when compared
to the optimal model (see Figure \ref{fig:bimodal}). Below we
investigate if this deviation from optimality can be related to the cue
combination strategy.

@@

A sample of \(N =\) 100 participants was recruited online through Amazon
Mechanical Turk. We used the same exclusion criteria as in Experiment 1.
\(N =\) 7 participants were excluded because they had less than 50\%
accurate responses on the unambiguous training trials. The final sample
consisted of \(N =\) 93 participants.

@@

\hypertarget{ref-barnhart2018}{}
Barnhart, W. R., Rivera, S., \& Robinson, C. W. (2018). Different
patterns of modality dominance across development. \emph{Acta
Psychologica}, \emph{182}.
